---
title: "Assng2_Stats"
output: github_document
author:
- name: Zuhaa 
- name: Saira 
- name: Lisa
date: "2023-03-01"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("lars")
library(lars)
library(tidyverse)
#install.packages("glmnet")
library(glmnet)
#install.packages("leaps")
library(leaps)
#install.packages("boot")
library(boot)
```

## Problem 1 
#(1) In this exercise, we will generate simulated data, and will then use this data to perform best subset selection.
#(a) Use the rnorm() function to generate a predictor X of length n = 100, as well as a noise vector e of length n = 100.
```{r}
set.seed(400)
X <- rnorm(100)
e <- rnorm(100)

```

#(b) Generate a response vector Y of length n = 100 according to the model ... are constants of your choice.
```{r}
Y <- 1+2*X + 3*X^2 - 4*X^3 + e
```

#(c) Use the regsubsets() function to perform best subset selection in order to choose the best model containing the predictors X,X2, . . . ,X10. What is the best model obtained according to AIC, BIC, and adjusted R2? Show some plots to provide evidence for your answer, and report the coefficients of the best model obtained. Note you will need to use the data.frame() function to create a single data set containing both X and Y.

```{r}
#Creating a data-frame using Y and X
df= data.frame(Y, X)

#Using regsubset function to determine the best, up to ten variable models
mod <- regsubsets(Y~poly(X,10), data=df, nvmax=10) 
(mod_summ = summary(mod))

# Set up a 2x2 grid so we can look at 4 plots at once
par(mfrow = c(2,2))

#plotting the Residual Sums of Squares for each of the ten polynomial models.
plot(mod_summ$rss, xlab = "Number of Variables", ylab = "RSS", type = "l")
#Determines the minimum RSS value from all models - 10
rss_min = which.min(mod_summ$rss)
#Adding the min value to the plot
points(rss_min, mod_summ$rss[rss_min], col ="red", cex = 2, pch = 20)


#plotting the Adjusted R^2 for each of the ten polynomial models.
plot(mod_summ$adjr2, xlab = "Number of Variables", ylab = "Adjusted RSq", type = "l")
#Determines the maximum Adjusted R^2 value from all models - 5 
adj_r2_max = which.max(mod_summ$adjr2)
#Adding the max value to the plot
points(adj_r2_max, mod_summ$adjr2[adj_r2_max], col ="red", cex = 2, pch = 20)

#plotting the Cp for each of the ten polynomial models.
plot(mod_summ$cp, xlab = "Number of Variables", ylab = "Cp", type = "l")
#Determining the minimum Cp values across all ten models - 4
cp_min = which.min(mod_summ$cp) 
#Adding the min value to the plot
points(cp_min, mod_summ$cp[cp_min], col = "red", cex = 2, pch = 20)

#plotting the BIC for each of the ten polynomial models.
plot(mod_summ$bic, xlab = "Number of Variables", ylab = "BIC", type = "l")
#Determining the minimum BIN values across all ten models - 3
bic_min = which.min(mod_summ$bic)
#Adding the min value to the plot
points(bic_min, mod_summ$bic[bic_min], col = "red", cex = 2, pch = 20)

```

From this model we see that the predictor variable X, by itself, is always significant. It is always included in the model. From the different metrics, we get slightly different best-fit models. The model with the lowest Cp score included 4 predictor (X^1, X^2, X^3, X^10), while the lowest BIC score only included 3 (X^1, X^2, X^3). Additional, the model with the highest adjusted R-squared included 5 predictors (X^1, X^2, X^3, X^6, X^10). 
Given that really all metrics agree that the first three preditors need be included (X^1, X^2, X^3)
From the plots of various metrics of model aqeqancy we see that the best model is one with three predictor vairables. from the summary output we see that this is the model 

```{r}
coef(mod, 3)
```

# (d) 

**Repeat (c), using forward stepwise selection and also using backwards stepwise selection. How does your answer compare to the results in (c)?**


```{r}
# Forward
regfit_fwd = regsubsets(Y~poly(X,10), data=df, nvmax = 10, method = "forward")
summary(regfit_fwd)
```


```{r}
# Backward
regfit_bwd = regsubsets(Y~poly(X,10), data=df, nvmax = 10, method = "backward")
summary(regfit_bwd)
coef(regfit_bwd, 3)
```



## Q2 

**We will now perform cross-validation on a simulated data set**
a) Generate a simulated data set as follows: 

```{r}
set.seed (1)
x <- rnorm (100)
y <- x - 2 * x^2 + rnorm (100)
```

#In this data set, what is n and what is p? Write out the model used to generate the data in equation form.
```{r}
#The n is 100, given that there are 100 observed (y) values. The p is 2 since there is only one predictor variable (x).
#$y_i$ = $\beta_0$ + $\beta_1$$x$ - $\beta_2$$x^2$ + $\epsilon_i$

```

**(b) Create a scatterplot of X against Y . Comment on what you find.**


```{r}
plot(x, y)
#from the plot we see that the relationship between x and y appears more quadratic then linear. 
#X range from -2 to 2, and Y from -8 to 2 

```

**c) Set a random seed, and then compute the LOOCV errors that result from fitting the following four models using least squares:**
#Note you may find it helpful to use the data.frame() function to create a single data set containing both X and Y.

```{r}
#i.
df2 = data.frame(y,x)

set.seed(101)
glm.fit = glm(y~x)
cv.glm(df2,glm.fit)$delta[1]

#ii.
fit.glm.2 <- glm(y~ poly(x,2))
cv.glm(df2, fit.glm.2)$delta[1]

#iii.
fit.glm.3 <- glm(y~poly(x,3))
cv.glm(df2, fit.glm.3)$delta[1]

#iv.
fit.glm.4 <- glm(y~poly(x,4))
cv.glm(df2, fit.glm.4)$delta[1]

```

#(d) Repeat (c) using another random seed and report your results. Are your results the same as what you got in (c)? Why?
```{r}
set.seed(102)

#i.
glm.fit = glm(y~x)
cv.glm(df2,glm.fit)$delta[1]

#ii.
fit.glm.2 <- glm(y~ poly(x,2))
cv.glm(df2, fit.glm.2)$delta[1]

#iii.
fit.glm.3 <- glm(y~poly(x,3))
cv.glm(df2, fit.glm.3)$delta[1]

#iv.
fit.glm.4 <- glm(y~poly(x,4))
cv.glm(df2, fit.glm.4)$delta[1]

#The results after changing the random seed are the same since LOOCV still evaluated all n folds containing a single observation regardless of the seed.
```

#(e) Which of the models in (c) had the smallest LOOCV error? Is this what you expected? Explain your answer.
```{r}
#According to the plot from 2b., we expected the true form of Y is quadratic. Therefore, it is expected that the lowest LOOCV test error rate is the quadratic polynomial. 
```

#(f) Comment on the statistical significance of the coefficient estimates that results from fitting each of the models in (c) using least squares. Do these results agree with the conclusions drawn based on the cross-validation results?
```{r}
summary(fit.glm.4)
#The p-values show that linear and quadratic values are statistically significant while cubic and 4th degree values are not. This result is consistent with the cross validation results above. 
```

