---
title: "Assng2_Stats"
output: github_document
author:
- name: Zuhaa 
- name: Saira 
- name: Lisa
date: "2023-03-01"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("lars")
library(lars)
library(tidyverse)
#install.packages("glmnet")
library(glmnet)
#install.packages("leaps")
library(leaps)
#install.packages("boot")
library(boot)
```

## Problem 1 
#(1) In this exercise, we will generate simulated data, and will then use this data to perform best subset selection.
#(a) Use the rnorm() function to generate a predictor X of length n = 100, as well as a noise vector e of length n = 100.
```{r}
set.seed(400)
X <- rnorm(100)
e <- rnorm(100)

```

#(b) Generate a response vector Y of length n = 100 according to the model ... are constants of your choice.
```{r}
Y <- 1+2*X + 3*X^2 - 4*X^3 + e
```

#(c) Use the regsubsets() function to perform best subset selection in order to choose the best model containing the predictors X,X2, . . . ,X10. What is the best model obtained according to AIC, BIC, and adjusted R2? Show some plots to provide evidence for your answer, and report the coefficients of the best model obtained. Note you will need to use the data.frame() function to create a single data set containing both X and Y.

```{r}
#Creating a data-frame using Y and X
df= data.frame(Y, X)

#Using regsubset function to determine the best, up to ten variable models
mod <- regsubsets(Y~poly(X,10), data=df, nvmax=10) 
(mod_summ = summary(mod))

# Set up a 2x2 grid so we can look at 4 plots at once
par(mfrow = c(2,2))

#plotting the Adjusted R^2 for each of the ten polynomial models.
plot(mod_summ$adjr2, xlab = "Number of Variables", ylab = "Adjusted RSq", type = "l")
#Determines the maximum Adjusted R^2 value from all models - 5 
adj_r2_max = which.max(mod_summ$adjr2)
#Adding the max value to the plot
points(adj_r2_max, mod_summ$adjr2[adj_r2_max], col ="red", cex = 2, pch = 20)

#plotting the Cp for each of the ten polynomial models. Cp will give the same reuslts as the AIC value.
plot(mod_summ$cp, xlab = "Number of Variables", ylab = "Cp", type = "l")
#Determining the minimum Cp values across all ten models - 4
cp_min = which.min(mod_summ$cp) 
#Adding the min value to the plot
points(cp_min, mod_summ$cp[cp_min], col = "red", cex = 2, pch = 20)

#plotting the BIC for each of the ten polynomial models.
plot(mod_summ$bic, xlab = "Number of Variables", ylab = "BIC", type = "l")
#Determining the minimum BIN values across all ten models - 3
bic_min = which.min(mod_summ$bic)
#Adding the min value to the plot
points(bic_min, mod_summ$bic[bic_min], col = "red", cex = 2, pch = 20)

```

From the different metrics (adjR^2, Cp/AIC, BIC), we get slightly different best-fit models. The model with the lowest Cp/AIC score included 4 predictors (X^1, X^2, X^3, X^10), while the lowest BIC score only included 3 (X^1, X^2, X^3). Additionally, the model with the highest adjusted R-squared included 5 predictors (X^1, X^2, X^3, X^6, X^10). All metrics agree that the first three predictors need to be included (X^1, X^2, X^3), whereas some suggest additional variables. BIC, in particular, tends to penalize large models, which explain the use of only the minimum number of predictors. Overall, the best fit models all include at least 3 polynomials (X^1, X^2, X^3).

# (d) 

**Repeat (c), using forward stepwise selection and also using backwards stepwise selection. How does your answer compare to the results in (c)?**


```{r}
# Forward selection using regsubset function
regfit_fwd = regsubsets(Y~poly(X,10), data=df, nvmax = 10, method = "forward")
fwd_summ = summary(regfit_fwd)

# Set up a 2x2 grid so we can look at 4 plots at once
par(mfrow = c(2,2))

#plotting the Adjusted R^2 for each of the ten polynomial models.
plot(fwd_summ$adjr2, xlab = "Number of Variables", ylab = "Adjusted RSq", type = "l")
#Determines the maximum Adjusted R^2 value from all models - 5 
adj_r2_max_fwd = which.max(fwd_summ$adjr2)
#Adding the max value to the plot
points(adj_r2_max_fwd, fwd_summ$adjr2[adj_r2_max_fwd], col ="red", cex = 2, pch = 20)

#plotting the Cp for each of the ten polynomial models. Cp will give the same reuslts as the AIC value.
plot(fwd_summ$cp, xlab = "Number of Variables", ylab = "Cp", type = "l")
#Determining the minimum Cp values across all ten models - 4
cp_min_fwd = which.min(fwd_summ$cp) 
#Adding the min value to the plot
points(cp_min_fwd, fwd_summ$cp[cp_min_fwd], col = "red", cex = 2, pch = 20)

#plotting the BIC for each of the ten polynomial models.
plot(fwd_summ$bic, xlab = "Number of Variables", ylab = "BIC", type = "l")
#Determining the minimum BIN values across all ten models - 3
bic_min_fwd = which.min(fwd_summ$bic)
#Adding the min value to the plot
points(bic_min_fwd, fwd_summ$bic[bic_min_fwd], col = "red", cex = 2, pch = 20)


```
The minimum and maximum Cp/AIC, BIC, and adj R^2 values for the model show the same results as the in c). Forward selection also suggests that at least three predictors are necessary (X^1, X^2, X^3). 

```{r}
# Backward selection using regsubset function 
regfit_bwd = regsubsets(Y~poly(X,10), data=df, nvmax = 10, method = "backward")
bwd_summ = summary(regfit_bwd)
# Set up a 2x2 grid so we can look at 4 plots at once
par(mfrow = c(2,2))

#plotting the Adjusted R^2 for each of the ten polynomial models.
plot(bwd_summ$adjr2, xlab = "Number of Variables", ylab = "Adjusted RSq", type = "l")
#Determines the maximum Adjusted R^2 value from all models - 5 
adj_r2_max_bwd = which.max(bwd_summ$adjr2)
#Adding the max value to the plot
points(adj_r2_max_bwd, bwd_summ$adjr2[adj_r2_max_bwd], col ="red", cex = 2, pch = 20)

#plotting the Cp for each of the ten polynomial models. Cp will give the same reuslts as the AIC value.
plot(bwd_summ$cp, xlab = "Number of Variables", ylab = "Cp", type = "l")
#Determining the minimum Cp values across all ten models - 4
cp_min_bwd = which.min(bwd_summ$cp) 
#Adding the min value to the plot
points(cp_min_bwd, bwd_summ$cp[cp_min_bwd], col = "red", cex = 2, pch = 20)

#plotting the BIC for each of the ten polynomial models.
plot(bwd_summ$bic, xlab = "Number of Variables", ylab = "BIC", type = "l")
#Determining the minimum BIN values across all ten models - 3
bic_min_bwd = which.min(bwd_summ$bic)
#Adding the min value to the plot
points(bic_min_bwd, bwd_summ$bic[bic_min_bwd], col = "red", cex = 2, pch = 20)


```

Similarly, backward selection also produces the same Cp/AIC, BIC and adj R^2 model results as the previous two sub-setting methods. Each metric suggests keeping in the first three predictors (X^1, X^2, X^3), whereas some additional predictors (X^6 or X^10) are included to reduce Cp/AIC and improve Adj R^2.

Overall, the results from the three methods are all very similar and suggest the same predictors for the best - fit model. 

## Q2 

**We will now perform cross-validation on a simulated data set**
a) Generate a simulated data set as follows: 

```{r}
set.seed (1)
x <- rnorm (100)
y <- x - 2 * x^2 + rnorm (100)
```

#In this data set, what is n and what is p? Write out the model used to generate the data in equation form.
```{r}
#The n is 100, given that there are 100 observed (y) values. The p is 2 since there is only one predictor variable (x).
#$y_i$ = $\beta_0$ + $\beta_1$$x$ - $\beta_2$$x^2$ + $\epsilon_i$

```

**(b) Create a scatterplot of X against Y . Comment on what you find.**


```{r}
#Adding the plot
plot(x, y)
#from the plot we see that the relationship between x and y appears more quadratic then linear. 
#X range from -2 to 2, and Y from -8 to 2 

```

**c) Set a random seed, and then compute the LOOCV errors that result from fitting the following four models using least squares:**
#Note you may find it helpful to use the data.frame() function to create a single data set containing both X and Y.

```{r}
#Creating a dataframe
df2 = data.frame(y,x)
set.seed(101)

#i.Linear model Y = beta_0 + beta_1*X + epsilon 
glm.fit = glm(y~x)
#Generate LOOCV errors
cv.glm(df2,glm.fit)$delta[1]

#ii. Quadratic model Y= beta_0 + beta_1*X + beta_2*^2 + epsilon 
fit.glm.2 <- glm(y~ poly(x,2))
#Generate LOOCV errors
cv.glm(df2, fit.glm.2)$delta[1]

#iii.Cubic model Y= beta_0 + beta_1*X + beta_2*^2 + beta_3*X^3 + epsilon 
fit.glm.3 <- glm(y~poly(x,3))
#Generate LOOCV errors
cv.glm(df2, fit.glm.3)$delta[1]

#iv. 4th degree model Y= beta_0 + beta_1*X + beta_2*^2 + beta_3*X^3 + beta_4*X^4 + epsilon 
fit.glm.4 <- glm(y~poly(x,4))
#Generate LOOCV errors
cv.glm(df2, fit.glm.4)$delta[1]

```

#(d) Repeat (c) using another random seed and report your results. Are your results the same as what you got in (c)? Why?
```{r}
set.seed(102)

#i.Linear model Y = beta_0 + beta_1*X + epsilon
glm.fit = glm(y~x)
#Generate LOOCV errors
cv.glm(df2,glm.fit)$delta[1]

#ii. Quadratic model Y= beta_0 + beta_1*X + beta_2*^2 + epsilon 
fit.glm.2 <- glm(y~ poly(x,2))
#Generate LOOCV errors
cv.glm(df2, fit.glm.2)$delta[1]

#iii.Cubic model Y= beta_0 + beta_1*X + beta_2*^2 + beta_3*X^3 + epsilon 
fit.glm.3 <- glm(y~poly(x,3))
#Generate LOOCV errors
cv.glm(df2, fit.glm.3)$delta[1]

#iv. 4th degree model Y= beta_0 + beta_1*X + beta_2*^2 + beta_3*X^3 + beta_4*X^4 + epsilon 
fit.glm.4 <- glm(y~poly(x,4))
#Generate LOOCV errors
cv.glm(df2, fit.glm.4)$delta[1]

#The results after changing the random seed are the same since LOOCV still evaluated all n folds containing a single observation regardless of the seed.
```

#(e) Which of the models in (c) had the smallest LOOCV error? Is this what you expected? Explain your answer.
```{r}
#According to the plot from 2b., we expected the true form of Y is quadratic. Therefore, it is expected that the lowest LOOCV test error rate is the quadratic polynomial. 
```

#(f) Comment on the statistical significance of the coefficient estimates that results from fitting each of the models in (c) using least squares. Do these results agree with the conclusions drawn based on the cross-validation results?
```{r}
summary(fit.glm.4)
#The p-values show that linear and quadratic values are statistically significant while cubic and 4th degree values are not. This result is consistent with the cross validation results above. 
```

